---
title: "Project 1 -- Grade Analysis of Two High Schools"
author: "Hui Zhang"
date: "Tuesday, October 21, 2014"
output: html_document
---
***
##Introduction 
```{r,echo=FALSE}
grade <- read.table(file="Grade_Analysis.txt", header=T)

```
This is the project for the first half of the semester for course Microbiol 612.This project will analyze the final grades of first-year students from two different high schools. In this project, we will analyze the difference or similarity between them. All the data are download from [wenku.baidu.com](http://wenku.baidu.com/view/09f6dc3752ea551811a68727.html).      

The project focuses on two high schools. These two high schools are the Top 2 in the district. There are `r summary(grade$School)["D"]` students in D shool that participated in this final exam, and `r summary(grade$School)["E"]` students in E school that participated in the exam. In sum,  `r nrow(grade)` students took the exam. In this final exam, there are `r ncol(grade)-2` subjects: `r colnames(grade)[3:10]`.           
      
* Are the two high schools really different from each other? Both schools are good, but some parents and students still want to know which one is the best. And then students will take the best school as their goal.      
* Is that true that students have a higher grade in Chinese than English? It seems like to be true in that Chinese is the native language to the test students. However, the exam for Chinese is much more complicated than English. Chinese exam involving more comphrehensive problems. So it's hard to get the conclusion.         
* Is there a positive correlation between Math and Physics? Physics might be the most difficult subject for first-year high students. Not all the students can handle Physics at the beginning. However, when I was in high school, students in my class, who were good at Physics, also did well at Math.      
* Which subject contributes most to the total grade? If a students do well in all the subjects, he will definitely get the high total grade. But it happens that he could get a high total grade even he has some weakness in certain subject.       

Questions in this project might not be such important, but these questions confused me when I was young. In fact, I also wonder is it true that girls do well in Arts while boys do well in Science. But I cannot get the data of sex of the students. Therefore, I do this grade analysis.


##Data Analysis    

* ###Basic analysis of total grades in both schools    
```{r,echo=FALSE}
school.total.mean <- aggregate(grade$Total,by=list(grade$School),mean)
school.total.sd <- aggregate(grade$Total,by=list(grade$School),sd)
school.total.max <- aggregate(grade$Total,by=list(grade$School),max)
school.total.min <- aggregate(grade$Total,by=list(grade$School),min)

```
The maximum total grades in D school and E school are `r school.total.max$x`. The minimum total grades in D school and E school are `r school.total.min$x`. The mean total grades in D school and E school are `r school.total.mean$x`. The standard deviation for total grades of D school and E school are `r school.total.sd$x`. These summary information can be plotted out as below:      
```{r,echo=FALSE}
boxplot(grade$Total~grade$School,xlab="School",ylab="Total Grade",main="Summary of Total Grade")

```
       
From the figure above, we can see that the average total grade of E school is a little higher than that of D school. And the total grade of E school is more centralized, while D school is more dispersed. ***Then, do the total grades of D and E school follow normal distribution?***      
```{r,echo=FALSE}
par(mfrow = c(1, 1))
d.total.hist <- hist(grade$Total[grade$School=="D"],plot=F,breaks=200)
plot(d.total.hist$density~d.total.hist$mids,type="h",ylab="Density",xlab="Total Grade",main="D School Total Grade Distribution")
d.total.mean <- mean(grade$Total[grade$School=="D"])
d.total.sd <- sd(grade$Total[grade$School=="D"])
points(0:1500,dnorm(0:1500,d.total.mean,d.total.sd),type="l",col="red",lwd=2)

e.total.hist <- hist(grade$Total[grade$School=="E"],plot=F,breaks=300)
plot(e.total.hist$density~e.total.hist$mids,type="h",ylab="Density",xlab="Total Grade",main="E School Total Grade Distribution")
e.total.mean <- mean(grade$Total[grade$School=="E"])
e.total.sd <- sd(grade$Total[grade$School=="E"])
points(0:1500,dnorm(0:1500,e.total.mean,e.total.sd),type="l",col="red",lwd=2)

```     
        
From the figures above, we can see that, total grade of E school follows normal distribution well, while D school are not well normally distributed. And the distribution of total grade of D school are relatively more dispersed. This conclusion is in accordance with the Summary figure above. But we can still take D school as normal distribution in future analysis.       
From figure of Summary, we can see that the average grade is a little different. To be more precise, the mean total grade of D school and E school are `r school.total.mean$x`. ***Then, are the total grades of these two school differ from each other?***  Let's check it with `t test`.    
Null hypothesis (H0) is: no difference between the two means.      
And the alternate hypothesis (H1) is: two samples are different from each other.    
Then, the result of `Welch Two Sample t-test` is:        
```{r,echo=FALSE}
d.total <- grade$Total[grade$School=="D"]
e.total <- grade$Total[grade$School=="E"]
t.test(d.total,e.total)

```
P value of `Welch Two Sample t-test` is `r t.test(d.total,e.total)$p.value`, which is much smaller than 0.05. The probability of `Type I Error` is very small. It means, our null hypothesis should be rejected. Therefore, there is a siginificant difference between D and E school.        
***Then, is that reliable that the mean total grade of E school is greater than that of D school?*** We still use `Welch Two Sample t-test`.    
Our null hypothesis (H0) is: E school is same as D school in total grade,       
And the alternative hypothesis (H1) is: E school is better than D school. The result is：    
```{r，echo=FALSE}
t.test(e.total,d.total,alternative="greater")

```
     
P value of `Welch Two Sample t-test` is `r t.test(e.total,d.total,alternative="greater")$p.value`, which is much smaller than 0.05. It means probility of `Type I Error` is very small. As a result, our null hypothesis should be rejected, and alternative hypothesis might be true. Therefore, in this final exam, E school did better than D school.   
      

* ###Relationship between separate subjects               

All the following analysis between subjects are based on data of E school.           

+ ####Chinese and English          

***It is reasonable to assume that students do better in Chinese than English, in that Chinese is the native language for the students. Is that true?***            
Here, null hypothesis (H0) is: There is no difference between Chinese and English.     
And alternative hypothesis (H1) is: Chinese grade is higher than English. Then,
```{r,echo=FALSE}
e.chinese.p <- grade$Chinese[grade$School=="E"]/150
e.eng.p <- grade$English[grade$School=="E"]/150
t.test(e.chinese.p,e.eng.p,alternative="greater")

```
From `t.test` above, P value is `r t.test(e.chinese.p,e.eng.p,alternative="greater")$p.value`, which is very high. It means probability of Type I Error is high. So we cannot reject null hypothesis. But it does not mean we can accept null hypothesis. In fact, it only suggests that there is not sufficient evidence against null hypothesis in favour of alternative hypothesis. Therefore, we only know grade of Chinese and English are different.       

+ ####Math and Physics    
```{r,echo=FALSE}
par(mfrow = c(1, 1))
e.math <- grade$Math[grade$School=="E"]
e.phys <- grade$Physics[grade$School=="E"]
plot(e.phys~e.math,pch=20,xlab="Math Grade",ylab="Physics Grade",main="E School Math Grade vs Physics Grade",xlim=c(0,150),ylim=c(0,100))
segments(x0=0,x1=150*0.6,y0=60,y1=60,lty=2,col="green",lwd=2)
segments(x0=150*0.6,x1=150*0.6,y0=0,y1=60,lty=2,col="green",lwd=2)
segments(x0=0,x1=150*0.95,y0=95,y1=95,lty=2,col="blue",lwd=2)
segments(x0=150*0.95,x1=150*0.95,y0=0,y1=95,lty=2,col="blue",lwd=2)
legend(5,90,legend=c("Pass line","Excellent line"),col=c("green","blue"),lwd=2,lty=2)

```
        
Green lines in the figure above, means 60% of the maximum possible grade, i.e. the pass grade. And blue llines means 95% of the maximum possible grade, i.e. the excellent grade. We can see that, most students passed the exam of Math and Physics. But only a few students get the excellent grade.     

When solving physical problems, it is inevitable to use mathematical skills. ***Then, can we get a conclusion that higher Physics grade depends on higher Math grade?*** We can test this by testing if there exist a positive linear correlation between Math grade and Physics grade.         
***Are Math and Physics grade normally distributed?*** We need to check this first.    
```{r,echo=FALSE}
e.math.hist <- hist(grade$Math[grade$School=="E"],plot=F,breaks=200)
plot(e.math.hist$density~e.math.hist$mids,type="h",ylab="Density",xlab="Math Grade",main="E School Math Grade Distribution")
e.math.mean <- mean(e.math)
e.math.sd <- sd(e.math)
points(0:1500,dnorm(0:1500,e.math.mean,e.math.sd),type="l",col="red",lwd=2)
```

```{r,echo=FALSE}
e.phys.hist <- hist(e.phys,plot=F,breaks=200)
plot(e.phys.hist$density~e.phys.hist$mids,type="h",ylab="Density",xlab="Physics Grade",main="E School Physics Grade Distribution")
e.phys.mean <- mean(e.phys)
e.phys.sd <- sd(e.phys)
points(0:1500,dnorm(0:1500,e.phys.mean,e.phys.sd),type="l",col="red",lwd=2)
```
       
From figures above, we can see that Math grade is well normally distributed. But Physics grade is roughly normal distributed. Then, we can use `Pearson's product-moment correlation` test.    
If the alternative hypothesis (H1) is: the grade of Physics is positively associated with the grade of Math. Then the test result is:

```{r,echo=FALSE}
cor.test(e.phys,e.math,method="pearson",alternative="greater")

```
From `Pearson's product-moment correlation` test, we get P value equals `r  cor.test(e.phys,e.math,method="pearson",alternative="greater")$p.value`. This small P value means we should reject our null hypothesis, and the alternative hypothesis might be true. And the `Pearson product-moment correlation coefficient` is 0.669. Therefore, there might be a positive correlation between Math and Physics grade.         
Then we can find is there a linear regression between the two subjects.      
```{r,echo=FALSE}
plot(e.phys~e.math,pch=20,xlab="Math Grade",ylab="Physics Grade",main="E School Math Grade vs Physics Grade",xlim=c(0,150),ylim=c(0,100))
legend(5,95,legend=c("Linear model"),col=c("red"),lwd=2,lty=1)
abline(b=0.6369,a=-4.9655,lwd=2,col="red")
summary(lm(e.phys~e.math))

```                 
       
From the analysis above, we know that there exists a linear regression, with intercept of -4.9655 and slope of 0.6369. From the figure above, we can see that students with higher Physics grade also have higher Math grade. And there are still some students who have higher Math grade, but relative low Physics grade. Therefore, good Math skill does contribute to higher Physics, but it is not a rule. 


* ###Relationship between subject and total grade       

All the following analysis are based on data from E school.        
Though all the subjects contribute to the total grade, they might have different weight. I am not going to analyze correlation betwwen all the subjects and total grade. I'd like to pick up two subjects, Physics and Histroy. These two subjects have the same maximum possible grade, 100. ***Can we assume that Physics has greater impact on total grades?*** Here, we could test the weight of correlation for two subjects, Physics and History.        

+ ####Physics, History vs Total grade          

We still use `Pearson's product-moment correlation` test.    
For Physics:        

```{r,echo=FALSE}
cor.test(e.total,e.phys,method="pearson")

```
The `Pearson product-moment correlation coefficient` is 0.6923.       

For History: 

```{r,echo=FALSE}
e.hist <- grade$History[grade$School=="E"]
cor.test(e.total,e.hist,method="pearson")

```
       
The `Pearson product-moment correlation coefficient` is 0.5402, which is smaller than the coefficient of Physics and total grade. Larger coefficient means closer correlation. Therefore, we verify the early assumption, that Physics contributes more to total grade.
     
##Conclusion      
Students in E high school might get higher grades than students in D high school. But we cannot say E school is much better than D school based on only one final exam. So when choosing high school to attend, it does not matter much which one you choose. So it would be more convenient to choose the one that is close to where you live. Students could do better in Chinese than English. This might be owing to Chinese is the native language for these students.  Math does help in Physics. But it does not mean if you have high Math grade, then you must get a high Physics grade. And it's not easy to get a good Physics grade without a good sense in Math. Subjects are not equally correlated to total grade. And Physics and total grade are more correlated than History and total grade.