---
title: "Project1"
author: "Hui Zhang"
date: "Saturday, October 18, 2014"
output: html_document
---


Introduction

This is the project for the first half of the semester for course Microbiol 612.This project will analyze Gross Output Value of Agriculture from 1949 to 2013 of Shanghai, China. All the data collected in this project is from [Shanghai Statistical Bureau](www.stats-sh.gov.cn).    

Shanghai is the largest city of by population density, as 3,800 per square kilometers, not only in China, but also all around the world. As a result, a large amount of food are required for its development. However, Shanghai is famous as a global financial center, and therefore, most analyses focuse on economy of Shanghai, but neglect agriculture of the city. This project collects data on agriculture from establishment of the People's Repulic of China, 1949, to 2013. And try to find out the growing tendency of agriculture of Shanghai, of different categories.
    
    
    
Data Analysis  

```{r,echo=FALSE}
agri <- read.table(file="data.txt", header=T)
colnames(agri)[2] <- "Total"
colnames(agri)[5] <- "Animal"
colnames(agri)[7] <- "Service"
agri

```



1. Plots     

Total vs Year

```{r,echo=FALSE}
plot(agri$Total~agri$Year,xlab="Year",ylab="Value,100 million Yuan",main="Total Gross Output Value of Agriculture",type="l",lwd=2)
legend(1954.417,298.218,legend="Gross Output Value",lwd=2)

```

Different Categories vs Year     

```{r,echo=FALSE}
plot(agri$Planting~agri$Year,xlab="Year",ylab="Value,100 million Yuan",main="Different Categorial Gross Output Value of Agriculture",type="l",lwd=2,col="purple")
points(agri$Forestry~agri$Year,xlab="Year",ylab="Value,100 million Yuan",main="Different Categorial Gross Output Value of Agriculture",type="l",lwd=2,col="red")
points(agri$Animal~agri$Year,xlab="Year",ylab="Value,100 million Yuan",main="Different Categorial Gross Output Value of Agriculture",type="l",lwd=2,col="blue")
points(agri$Fishery~agri$Year,xlab="Year",ylab="Value,100 million Yuan",main="Different Categorial Gross Output Value of Agriculture",type="l",lwd=2,col="green")
points(agri$Service~agri$Year,xlab="Year",ylab="Value,100 million Yuan",main="Different Categorial Gross Output Value of Agriculture",type="l",lwd=2,col="orange")
legend(1953.953,160.6015,legend=c("Planting","Forestry","Animal Husbandry","Fishery","Service Industry for Agricalture"),col=c("purple","red","blue","green","orange"),lty=1,lwd=2)

```


From the figures above, we can see that the data for agricultural output value does not follow normal distribution. Therefore, we cannot use t test, since t test requires data follow normal distribution. Finally, we decided to use a non-parametric test, i.e. Wilcox test, to analyze the data.  




One goal of this project is to find the relationship between totoal agriculture gross output value and different categories.    
Here we make an assumption that 
